% =====================================================================
%  AI-Driven Gaming Addiction Screening System
%  PES University Capstone -- PW26_SJ_05  (UE23CS320B, Phase-2)
%
%  This paper describes the system AS ACTUALLY BUILT (native Android apps +
%  cloud Flask backend + calibrated classical-ML ensemble). Metrics reported
%  here are the honest, current numbers from model_metadata.json.
%
%  Compile:  pdflatex PROJECT_PAPER.tex   (twice, for the ToC/refs)
% =====================================================================
\documentclass[11pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{newtxtext}   % dark, readable Times text (replaces thin Computer Modern)
\usepackage{newtxmath}   % matching math font; keeps real subscripts in formulae
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{caption}
\usepackage[hidelinks]{hyperref}
\usepackage{microtype}   % char protrusion + font expansion: far fewer words run off the margin
\usepackage[strings]{underscore}  % printable underscores in text; subscripts still work in math

\setlength{\parskip}{0.45em}
\setlength{\parindent}{0pt}
\setlength{\emergencystretch}{3em}  % last-resort stretch so lines don't overflow the right margin

\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  columns=fullflexible,
  keepspaces=true,
  showstringspaces=false,
  xleftmargin=0.5em,
}

\title{\textbf{AI-Driven Gaming Addiction Screening System}\\
\large A Multimodal, Privacy-Conscious Mobile Platform for Early
Detection and Parental Awareness}

\author{
Khushee P Kiran (PES1UG23CS303) \and
Kanak Goyal (PES1UG23CS279) \and
Vidisha Murali (PES1UG23CS681) \and
Kaustubh Agarwal (PES1UG23CS291) \\[0.5em]
\textit{Under the guidance of Prof. Suresh Jamadagni} \\
\textit{Department of Computer Science and Engineering, PES University} \\
\textit{Capstone Project PW26\_SJ\_05 (UE23CS320B), January--May 2026}
}
\date{}

\begin{document}
\maketitle

% ---------------------------------------------------------------------
\begin{abstract}
\noindent
Excessive mobile gaming is increasingly linked to disrupted sleep, mood and
routine in adolescents, yet most existing tools track only raw screen time and
miss the behavioural and emotional patterns that distinguish healthy play from
problematic use. We present a deployed, multimodal screening platform that fuses
three signals collected during real gameplay --- objective play behaviour, in-game
chat toxicity, and voice emotion --- into a single calibrated risk band (Casual /
At-Risk / Addicted). The system comprises two native Android applications (a Child
monitoring app and a Parent dashboard) talking to a cloud Flask backend that serves
a weighted ensemble of three classical machine-learning models. We deliberately
favour calibrated, explainable classical models (Random Forest, Logistic Regression,
Gradient Boosting) with SHAP attributions over heavyweight deep models, prioritising
on-device privacy, deployability and honesty of evaluation. The platform is
production-deployed (HTTPS, token authentication, parental consent, a durable
managed database), ships a parent-feedback loop that converts real verdicts into
training labels, and provides a parent-to-child nudge channel so the parent can
intervene --- not merely observe. We report honest, calibrated metrics --- including a frank account
of where the chat model over-flags on realistic data --- and frame the output
explicitly as \emph{screening and awareness, not clinical diagnosis}.
\end{abstract}

\tableofcontents
\vspace{1em}

% =====================================================================
\section{Introduction}
% =====================================================================
Mobile gaming is now a daily activity for most adolescents. While gaming is not
inherently harmful, a minority of users develop patterns that the DSM-5 (Internet
Gaming Disorder) and ICD-11 (Gaming Disorder) characterise behaviourally: impaired
control over play, escalating priority of gaming over other activities, and
continuation despite negative consequences. Conventional parental-control tools
report \emph{how long} a phone was used but not \emph{how} the child played --- the
late-night sessions, the rapid re-logins, the rising frustration in chat and voice.

This project builds a system that captures those richer signals during gameplay and
turns them into an interpretable risk band that a parent can act on with
\emph{awareness rather than punishment}. Our key design stance is honesty and
deployability: every claim in this paper reflects the system as actually built and
running, and the models are chosen to be lightweight, explainable, and calibrated
rather than to maximise a headline accuracy number.

\subsection*{Contributions}
\begin{enumerate}[leftmargin=1.4em]
  \item A deployed three-tier platform: two native Android apps + a cloud Flask
        backend serving a multimodal ML ensemble.
  \item A behavioural feature pipeline that computes ten \emph{measured} features
        from real session history and ten derived psychometric proxies through a
        \emph{single shared function}, eliminating train/serve skew.
  \item An availability-weighted ensemble that re-normalises over only the
        modalities that produced data, with isotonic probability calibration and
        per-prediction SHAP explanations.
  \item Production engineering: token authentication, peppered PIN hashing,
        HTTPS-only transport, encrypted on-device token storage, parental consent,
        raw-audio deletion, and a durable managed database --- plus free-tier load
        engineering (pooled, validated connections to the serverless database,
        per-client rate limiting behind the reverse proxy, and caching/throttling on
        the polling hot paths) so the deployment survives real concurrent use.
  \item A parent-feedback loop that records real ``accurate / false-alarm''
        verdicts --- the only source of genuine labels in an otherwise
        synthetically-trained system --- enabling future retraining.
  \item A parent-to-child nudge channel (preset or custom messages, plus an automatic
        ``mind the language'' nudge on toxic chat) that turns the platform from passive
        monitoring into two-way guidance.
  \item A custom-keyboard (Input Method Editor) chat-capture path that reliably records
        the child's typed in-game text even in canvas/engine games (e.g.\ Roblox) the
        accessibility API cannot read --- without resorting to screen OCR --- while
        capturing only the child's own typing.
  \item Tamper resistance and resilient delivery: a self-restarting foreground service,
        a battery-optimisation exemption step (OEM power managers silently killing the
        monitor are the most common real-world failure), a parent-PIN lock on
        logout/settings with parent-role-only credential and data-deletion changes, a
        server-side heartbeat watchdog that alerts the parent when monitoring goes silent
        (uninstall / kill / offline), a live monitoring-health and ``playing right now''
        indicator on the parent dashboard, an offline retry queue so chat captured
        without connectivity is delivered later instead of lost, and real-time alerts
        pushed to \emph{every} guardian device in a family.
  \item An honest evaluation, including the realistic-stream precision of the chat
        model and the synthetic-data caveat on the voice model.
\end{enumerate}

% =====================================================================
\section{Related Work}
% =====================================================================
\textbf{AI-based monitoring of gaming disorder.} Reviews and cohort studies establish
that gaming disorder is multifactorial --- behavioural, emotional and
psychological~\cite{ergin2025,kaya2023} --- and that continuous, AI-based monitoring
detects early warning signs more reliably than self-report surveys or raw screen-time
tools~\cite{huang2024,jiang2024}. This motivates our continuous, multi-signal approach
and our behaviour-dominant weighting.

\textbf{Language and chat signals.} Short pieces of user text carry usable
addiction-risk and emotional signal~\cite{strojny2024}, and sentiment models perform
well on fast, informal, gaming-specific chat --- including multilingual
settings~\cite{kina2024}. This grounds our chat-toxicity channel and our future-work
direction of a multilingual/code-mixed model.

\textbf{Voice and multimodal fusion.} Combining speech with text yields stronger
emotional inference than any single stream~\cite{tyagi2024}, and acoustic cues (pitch,
energy, timbre) carry emotional information usable for behavioural
monitoring~\cite{schuller2013}. This supports our voice-emotion channel and the weighted
fusion of behaviour, chat and voice.

\textbf{On-device privacy and parental tools.} On-device processing supports
personalisation while keeping sensitive data on the phone~\cite{chakraborty2025}, and
prior work on parental-monitoring design stresses awareness and trust over
surveillance~\cite{iftikhar2021}; commercial tools such as Bark~\cite{bark2025} show
demand but are largely screen-time/keyword based. We position our system as
\emph{screening and guidance, not punishment}, with privacy-conscious handling and
explainable output.

% =====================================================================
\section{System Architecture with Explanation}
% =====================================================================
The platform follows a three-tier architecture (Figure-equivalent shown below):
a \textbf{data-collection tier} (Child app), a \textbf{serving tier} (Flask backend +
ML ensemble), and a \textbf{presentation tier} (Parent app).

\begin{lstlisting}
            CHILD APP (Android, Kotlin)                 PARENT APP (Android, Kotlin)
   +------------------------------------+        +-------------------------------+
   | UsageStatsManager  -> GameDetector |        | Risk band + SHAP "why"        |
   | AccessibilityService -> typed chat |        | 14-day trend, alerts feed     |
   | AudioRecord + Vosk  -> voice + STT  |        | Feedback: Accurate/False alarm|
   | Behavioural feature engine          |        | Weekly PDF, limits, settings  |
   +------------------+-----------------+        +---------------+---------------+
                      |  HTTPS (token auth)              ^/v  HTTPS (poll + nudge)
                      v                                          |
        +-------------------------------------------------------------------+
        |                 CLOUD BACKEND  (Flask, gunicorn)                  |
        |  Session API | Chat/Voice ingest | Prediction engine | Alerts     |
        |  Auth (signed tokens) | Consent | Retention purge | PDF report    |
        +------------------------------+------------------------------------+
                                       |
            +--------------------------+--------------------------+
            v                          v                          v
     BEHAVIOUR (40%)            CHAT TOXICITY (30%)          VOICE EMOTION (30%)
   RandomForest (3-class)     LogReg + TF-IDF             GradientBoosting (4-class)
   + isotonic calibration     + keyword lexicon           + valence-arousal fusion
   + SHAP explanations
            \_________________________ weighted ensemble ____________________/
                                       v
                     Risk score 0..1  ->  Casual / At-Risk / Addicted
                                       v
                          Managed Postgres (Neon) / SQLite
\end{lstlisting}

\subsection{Data-collection tier (Child app)}
The Child app runs always-on foreground services that:
\begin{itemize}[leftmargin=1.4em]
  \item detect the foreground gaming app via \texttt{UsageStatsManager} and a
        layered detector (curated allow-list $\rightarrow$ deny-list $\rightarrow$
        OS \texttt{CATEGORY\_GAME} + launchable check, with parent overrides), so
        \emph{any} game is monitorable, not a fixed list;
  \item capture typed in-game chat through a custom keyboard (Input Method Editor)
        \emph{and} an \texttt{AccessibilityService}, so even canvas/engine games
        (e.g.\ Roblox) whose chat box exposes no text node are covered --- without
        fragile screenshot OCR;
  \item record microphone audio (\texttt{AudioRecord}, 16~kHz mono) and run
        \textbf{Vosk} on-device speech-to-text, so spoken words are transcribed
        \emph{without raw audio leaving the device for transcription};
  \item compute behavioural metrics from real session history.
\end{itemize}

\subsection{Serving tier (Flask backend + ML ensemble)}
A Flask application (run under gunicorn) exposes a REST API for session lifecycle,
chat/voice/behaviour ingestion, prediction, alerts, dashboards and a parent-feedback
endpoint. It loads three serialised scikit-learn models and fuses them. The backend
is \textbf{dual-dialect} --- the identical SQL runs on SQLite (local) or PostgreSQL
(production) --- and is deployed on a managed host with a durable managed Postgres.

Because both hosting tiers are free (a 512\,MB single-instance web service and a
serverless Postgres that scales to zero when idle), the serving tier is engineered
for load rather than assumed to have headroom: database connections are
\emph{pooled and validated on borrow} (a connection the idle-scaled database silently
killed is detected by a one-round-trip probe and replaced, instead of failing the
request), rate limits are applied per real client IP behind the reverse proxy (not per
proxy, which would make all users share one quota), SHAP explanations are memoised and
the anomaly scan is throttled on the 30-second dashboard polling path, and the single
threaded worker is sized so the whole ML stack stays within memory. Concurrent
\emph{audio} analyses are additionally capped by a semaphore: pitch/MFCC extraction
allocates tens of megabytes per clip, and during live testing several voice segments
arriving together briefly exceeded the instance's memory --- the cap (plus periodic
worker recycling and single-threaded BLAS) bounds the peak instead of trusting it.

\subsection{Presentation tier (Parent app)}
The Parent app authenticates with a family code + PIN, shows the current risk band
with its SHAP-derived ``why this level'' factors, a 14-day trend, an alerts feed, a
weekly PDF report, and the \textbf{feedback controls} that let the parent mark each
alert as accurate or a false alarm. It also offers a \textbf{``Send a nudge''} control
--- a preset or custom message that appears as a notification on the child's phone.

\subsection{Privacy and security design (as built)}
\begin{itemize}[leftmargin=1.4em]
  \item \textbf{Transport:} the released apps ship a network-security configuration
        that forbids cleartext (HTTPS only); the backend is TLS-terminated.
  \item \textbf{Authentication:} signed bearer tokens (HMAC, \texttt{itsdangerous})
        issued at login; per-user authorisation (a parent can read only their own
        children), plus \emph{role} restrictions on top: changing the family PIN and
        deleting a child's data require a parent token, so a child's own token cannot
        quietly disable the safeguards that monitor them.
  \item \textbf{Credentials:} PINs are stored only as keyed (peppered) hashes; raw
        PINs are never persisted. Tokens are kept in Android
        \texttt{EncryptedSharedPreferences}.
  \item \textbf{Raw data:} microphone audio is deleted immediately after feature
        extraction (\texttt{KEEP\_AUDIO=0}); a retention job purges old raw events.
  \item \textbf{Consent:} the Child app gates monitoring behind an explicit consent
        screen; data deletion is parent-controlled.
\end{itemize}
We note honestly that, unlike a fully on-device design, chat text and voice features
\emph{are} sent to the cloud for inference; the mitigations above (TLS, token auth,
raw-audio deletion, consent, retention limits) define the actual privacy posture.

% =====================================================================
\section{Methodology}
% =====================================================================
\subsection{Behavioural feature engineering (train/serve aligned)}
The behaviour model consumes 20 features. \textbf{Ten objective features} are
\emph{measured} from the user's real session history (not simulated from elapsed
time): daily\_play\_time\_hours, weekly\_play\_time\_hours, sessions\_per\_day,
avg\_session\_duration\_min, late\_night\_play\_ratio, days\_played\_per\_week,
longest\_play\_streak\_days, binge\_sessions\_per\_week,
avg\_break\_between\_sessions\_min and rapid\_relogin\_ratio. Hour-of-day signals
(late-night play, and the sleep-impact view built on it) are evaluated in the
\emph{child's local timezone} --- the device reports its UTC offset, and the server
shifts stored timestamps accordingly --- so a cloud host running in UTC does not
mis-time the night of a child in another timezone.

\textbf{Ten psychometric proxies} (urge, loss-of-time-awareness, control-loss,
craving, tolerance, missed-sleep, fatigue, routine-disruption,
neglect-responsibilities, gaming-priority) are derived deterministically from the
objective ten. Crucially, this derivation lives in a \emph{single shared function}
imported by both the trainer and the live backend, so the model is trained and
served on the identical feature definition --- eliminating the train/serve skew that
would otherwise invalidate the deployed predictions.

\subsection{Datasets and honesty of data}
The behaviour and voice models are trained on \emph{synthetic / illustrative} data
with realistic class overlap; the chat model is trained on a public toxicity corpus.
We are explicit that synthetic training proves \emph{internal validity and pipeline
correctness}, not real-world accuracy. The parent-feedback loop is the designed path
to genuine labels.

\subsection{Model development}
\begin{itemize}[leftmargin=1.4em]
  \item \textbf{Behaviour:} \texttt{RandomForestClassifier} (3-class), features
        standard-scaled; an isotonic \texttt{CalibratedClassifierCV} is fitted on a
        held-out split so served probabilities are meaningful, while the raw forest
        is retained for SHAP and feature importances.
  \item \textbf{Chat:} \texttt{TfidfVectorizer} (1--2 grams) +
        \texttt{LogisticRegression}, combined via \texttt{max()} with a curated
        keyword-lexicon booster (with gaming hard-negatives removed so normal terms
        like ``nice kill'' are not flagged).
  \item \textbf{Voice:} \texttt{GradientBoostingClassifier} over 17 acoustic
        features (13 MFCCs, pitch mean/std, energy mean/std), fused with text
        valence and chat toxicity through a valence--arousal rule.
\end{itemize}

\subsection{Ensemble fusion}
Let $b,c,v$ be the behaviour, chat and voice risk components with \emph{fixed prior}
weights $w_b{=}0.40$, $w_c{=}0.30$, $w_v{=}0.30$. These are clinically-motivated priors,
\emph{not} values fitted to labelled outcomes. Behaviour carries the largest weight
because both DSM-5 (Internet Gaming Disorder)~\cite{dsm5} and ICD-11 (Gaming
Disorder)~\cite{icd11} define the condition \emph{behaviourally} --- impaired control, escalating priority of gaming over
other activities, continuation despite harm --- and behaviour is also the only
\emph{objective, measured} channel. Chat and voice are equal, secondary, corroborating
emotional-distress signals and are noisier, so they share the remaining weight. We
deliberately do not fit the weights, because the training data is synthetic and fitting
on it would be circular; calibrating them from real labels is future work (via the
feedback loop).

The fused score re-normalises over only the modalities $P\subseteq\{b,c,v\}$ that
actually produced data, then applies a game-genre risk multiplier $g$:
\[
\text{risk} \;=\; g \cdot \frac{\sum_{m\in P} w_m\, s_m}{\sum_{m\in P} w_m}.
\]
The score maps to bands \textbf{Casual} ($<0.33$), \textbf{At-Risk} ($0.33$--$0.67$) and
\textbf{Addicted} ($\geq 0.67$). These cut-points are even tertiles of the $0$--$1$ score
--- again a transparent prior, not a fitted cut-off --- and they align with the score's
own anchor points: a confidently-Casual prediction scores $\approx 0$, a
confidently-At-Risk one $\approx 0.5$ (the behaviour score is $P(\text{at-risk})\cdot 0.5
+ P(\text{addicted})\cdot 1.0$), and a confidently-Addicted one $\approx 1.0$, so each
``pure'' prediction falls inside its matching band. They are \emph{screening} bands for
awareness, not diagnostic cut-offs.

\paragraph{Fixed priors, dynamic result.}
The constants above never change at run time; what adapts to each session is the value
\emph{computed from} them, through three mechanisms:
\begin{itemize}[leftmargin=1.4em]
  \item \textbf{Availability re-weighting.} The weights are renormalised over only the
        modalities present, so an absent one contributes nothing (e.g.\ with no voice,
        behaviour and chat become $\approx$57\,\%/43\,\% for that session). This avoids a
        missing modality injecting a placeholder value and capping behaviour-only
        sessions below ``Addicted''.
  \item \textbf{Game-genre multiplier ($g$).} The final score is scaled by a per-genre
        risk factor (a competitive, chat-heavy game weighs more than a casual puzzle),
        read from a fixed lookup.
  \item \textbf{Observation mode.} For the first few sessions ($<3$) the band is capped
        below ``Addicted'' regardless of score, so the system never over-calls on sparse
        data.
\end{itemize}
Thus the same fixed priors yield a context-sensitive result; the priors and cut-points
themselves would only \emph{change} by being re-learned from real feedback labels, which
is future work.

\subsection{Evaluation strategy}
Each model is evaluated on a held-out split it never trained on. For the chat model
we additionally report a \emph{realistic, naturally-imbalanced} held-out set, because
balanced-split accuracy is misleading for a rare-positive problem. Probability
calibration is measured by the multiclass Brier score.

% =====================================================================
\section{Implementation (Salient Modules)}
% =====================================================================
\subsection{Game detection and session lifecycle}
\texttt{GameDetector} + \texttt{PassiveMonitorService} identify the foreground game
and open a session via \texttt{POST /api/session/start}; the session id scopes all
subsequent uploads. A server-side safety net auto-closes any session left open
beyond a threshold (e.g.\ from a lost end-event), dating it at the last recorded
activity, so a dropped network event never leaves a child shown as ``perpetually
playing''.

\subsection{Chat capture: custom keyboard, Accessibility + on-device STT}
Typed chat is captured along two complementary paths (both posted with
\texttt{source=keyboard}); spoken words are transcribed on-device by Vosk and posted
with \texttt{source=voice\_stt}. For apps that expose a standard Android text field,
\texttt{ChatAccessibilityService} reads the typed text. Many games, however, render
their chat box inside their own canvas/engine surface (e.g.\ Roblox), where no Android
text node exists and the accessibility path sees nothing. To close that gap
deterministically --- and without the unreliability of screen OCR (which cannot know
where each game draws its chat) --- the Child app ships its own \emph{Input Method
Editor} (a custom keyboard). Because every keystroke passes through the keyboard before
reaching the game, it captures exactly the child's own typed text in \emph{any} app,
canvas games included, and never other players' messages or game UI. Capture is gated to
monitored games with an active session (it records nothing in other apps such as banking
or search), and an 8-second server-side de-duplicate prevents the keyboard and
accessibility paths from logging the same line twice. The prediction engine feeds
\emph{typed} chat to the toxicity channel and routes spoken words through the
voice/valence fusion, so the two are never double-counted.

Capture is also resilient to connectivity loss: a chat line whose upload fails on a
network error is held in a small bounded queue on the device (encrypted preferences,
capped and aged out after a day) and retried automatically once the phone is back
online --- previously such lines were silently lost. Only \emph{network} failures are
queued, never lines the server received and rejected, so retries cannot create
duplicates beyond the server's own de-duplication window.

\subsection{Voice recording and emotion (valence--arousal fusion)}
\texttt{VoiceRecorderService} records PCM segments and extracts 17 acoustic features
(13 MFCCs, pitch mean/std, energy mean/std), which the backend classifies into
angry / excited / frustrated / neutral. The acoustic model reliably separates
\emph{arousal} (animated vs.\ quiet speech) but cannot judge \emph{valence} (positive
vs.\ negative): loud-but-calm speech, or a child's naturally high-pitched voice, can be
mis-read as ``angry''. The spoken \emph{words} supply the missing valence.

The system therefore fuses the two on a valence--arousal plane. \emph{Arousal} is taken
from the acoustic probability distribution, while \emph{valence} is a weighted blend of
three signals computed on the Vosk transcript of the same segment: the acoustic
distribution as a weak prior, a gaming-aware lexicon (positive \emph{and} negative
words), and the \emph{trained chat-toxicity model} --- which detects hostility beyond the
word list and contributes strong negative valence. The resulting
$(\text{valence}, \text{arousal})$ point is mapped back to one of the four emotions, and
then to a voice-risk weight. This corrects the child-pitch bias (high arousal $+$ neutral
words $\rightarrow$ ``excited'', not ``angry'') and lets genuinely toxic spoken language
pull the read toward distress even when the tone is calm. If no transcript is available,
the system falls back to the acoustic label alone. Before any of this, an RMS
\emph{silence floor} rejects near-silent segments outright: the pipeline gain-normalises
each clip so the energy features are microphone-gain-invariant, and without the floor
that normalisation would amplify room noise up to speech level and classify the noise as
an emotion. A silent segment is recorded as low-confidence neutral instead --- an interim
form of the voice-activity detection planned in Future Work.

This is where the voice and chat models cooperate: the chat-toxicity model is reused to
score the \emph{spoken} words inside the voice fusion. Crucially, those transcribed words
feed \emph{only} the voice channel through this fusion --- the typed-chat toxicity
channel explicitly excludes them (\texttt{source = voice\_stt}) --- so a single utterance
is never double-counted across two modalities. Foreground-service start,
microphone-permission checks and Android-14 restrictions are handled defensively so a
denied permission never crashes the app.

\subsection{Prediction, calibration and explainability}
On session end the backend computes the 20 features, runs the three models, fuses
them, and stores the prediction with its component scores and a presence flag per
modality. The served probability comes from the \emph{calibrated} behaviour model; a
SHAP~\cite{shap2017} \texttt{TreeExplainer} produces the top signed factors shown to
the parent. Because the voice pipeline (record $\rightarrow$ buffer $\rightarrow$
analyse $\rightarrow$ fuse) is slower than the behaviour and chat paths, a voice segment
can land \emph{after} a short session's end-of-session prediction; the server therefore
re-scores the session when a late voice event arrives, so the voice modality still counts
on brief sessions rather than being silently dropped.

\subsection{Anomaly detection}
Alongside the per-session risk band, the backend flags \emph{anomalies} --- a session
that deviates sharply from the child's own recent history (e.g.\ a sudden spike in
duration or late-night play). It computes a $z$-score of the current session against the
mean and standard deviation of the user's prior sessions and raises a flag when the
deviation is large. This catches unusual escalation that a static threshold on a single
session might miss, and complements the global risk bands with a personalised
``different from this child's normal'' signal.

\subsection{Parent feedback loop}
Each alert in the Parent app carries an ``\emph{Was this assessment accurate?}''
control. A verdict is stored via \texttt{POST /api/feedback} together with a snapshot
of the model's category and score at that moment, giving a growing set of real labels
for future retraining; an aggregate \emph{agreement rate} is exposed as a real-world
precision proxy and shown back to the parent at the top of the alerts screen
(``based on your $N$ verdicts, you've rated the model accurate $X\%$ of the time''),
closing the loop visibly rather than collecting labels into a void.

\subsection{Parent $\rightarrow$ child nudges}
Beyond monitoring, the parent can \emph{act}. A ``Send a nudge'' control posts a message
--- preset (e.g.\ \emph{take a break}, \emph{mind your language}) or custom --- via
\texttt{POST /api/parent/nudge}. The child's always-on service polls
\texttt{GET /api/child/nudges} and pops each up as a notification exactly once (marked
delivered server-side). When the chat model detects confidently-toxic language the
system also queues an automatic, rate-limited self-correction nudge to the child. A
daily time limit the parent sets is likewise delivered to the child as a nudge (rather
than surfacing only in the parent's own alert feed). This makes awareness \emph{two-way}
--- the parent intervenes gently rather than only watching, in keeping with the
project's ``guidance, not punishment'' goal.

\subsection{Real-time alerts and multi-guardian delivery}
Material events raise a parent alert in real time: the start of a gaming session,
confidently-toxic chat, an unusual-behaviour anomaly, and the tamper signals below.
Each alert carries a server-computed age, so the app shows accurate relative times
(``2h ago'') regardless of the timezone gap between phone and server. Beyond discrete
alerts, the dashboard header shows a \emph{live status strip}: whether the child is
playing \emph{right now} (game and minutes elapsed, from the open session) and whether
the monitoring app is healthy (time since its last heartbeat check-in) --- so a parent
sees ``playing Roblox $\cdot$ 23 min'' or ``no check-in for 47 min'' at a glance instead
of waiting for the next alert.
Alerts reach guardians by two complementary paths. Every guardian phone polls
\texttt{GET /api/alerts} ($\sim$60\,s), and for high-risk events the server also sends a
Firebase Cloud Messaging push. Because both parents may monitor the same family under the
shared family code, push is delivered to \emph{all} registered guardian devices --- not
just the last to sign in: device tokens are stored per device and dead tokens are pruned
on send. To keep the free-tier memory footprint low, the Firebase SDK is imported and
initialised \emph{lazily}, only on the first push.

\subsection{Tamper resistance and reliability}
Monitoring is designed to survive ordinary interference and to make interference
visible. The data-collection service runs in the foreground, restarts itself if the
child swipes the app away from recents, and resumes on device boot; its setup flow also
requests exemption from battery optimisation, because aggressive OEM power managers
silently killing background services is the most common \emph{accidental} way monitoring
dies in the field. Services are guarded against double-starts (boot receiver and app
launch can both start them), so restarts never stack duplicate capture loops. A child
cannot quietly disable monitoring: logging out and changing settings require the family
parent PIN, verified server-side so the PIN is never exposed to the child, and a logout
itself raises a parent alert; the family PIN can only be \emph{changed} with a
parent-role token, and data deletion is likewise parent-only --- a child's own
authenticated session cannot rewrite or erase its safeguards. Sign-out and sign-in are
treated symmetrically and honestly: a logout immediately clears the monitoring-health
status (the dashboard must not keep claiming ``monitoring active'' off a heartbeat that
still looks fresh) and suppresses the now-redundant ``gone silent'' watchdog alert,
while signing back in raises a ``monitoring is active again'' alert and turns the
health indicator green at once. Because Android gives an app no callback at its own uninstall,
removal is detected \emph{indirectly}: the Child app emits a periodic heartbeat, and the
server raises a single ``monitoring has gone silent'' alert when no heartbeat arrives for
a set interval --- a catch-all covering uninstall, force-stop, being killed, or going
offline --- re-arming once the app checks in again. To keep this from crying wolf, the
watchdog respects the child's local \emph{quiet hours} (reported via the heartbeat's
device-timezone offset): it does not raise the alert overnight, when a silent phone is
almost certainly just off or asleep, but still fires by day if the app is genuinely gone.
An \emph{optional} device-administrator mode goes further: once the family activates it,
the app cannot be casually uninstalled (Android forces deactivation first) and the parent
is alerted the \emph{instant} someone attempts to disable it (\texttt{onDisableRequested}
fires before removal completes). It is strictly opt-in and requests \emph{no} lock/wipe
policies --- only the uninstall block and the disable hook --- keeping it appropriate for
a consent-based model rather than a punitive lockdown.

\subsection{Security and accounts}
A family registers once and receives a unique family code; each child has a private
PIN, and the parent signs in with the family code + a shared PIN. All protected
endpoints enforce token authentication and per-user authorisation. The family code
remains permanently visible (and tap-to-copy) in the Child app's PIN-gated Settings
--- it is an identifier, not a credential (every use of it also requires the family
PIN) --- so a parent who didn't note it down at registration is never locked out of
the Parent app.

\subsection{Parental dashboard and wellbeing}
The Parent app shows the day's risk --- a \emph{duration-weighted} roll-up of that
day's sessions, so a brief session (e.g.\ a one-minute trial of a game) cannot dominate
the headline the way a single last-session score would --- together with its SHAP
factors, a 14-day trend, most- and recently-played games, alerts, anomaly flags, a
weekly PDF report, daily-limit controls and notification handling; the chat view
separates text the child \emph{typed} from speech that was transcribed. The dashboard
auto-refreshes in the background without disturbing the view (it re-renders only when the
data actually changes), and a parent can switch between children in the family without
re-entering the family code. The Child app adds wellbeing features --- a rule-based
``Mira'' check-in companion, a gamified daily check-in (mood / sleep / energy) with a
check-in streak, reflections, breathing resets and gentle break / limit nudges ---
supporting awareness rather than hard restriction.

% =====================================================================
\section{Results}
% =====================================================================
We report the honest, current metrics from the deployed models. Headline accuracy is
deliberately \emph{not} inflated; calibration and realistic-stream behaviour are
reported alongside.

\subsection{Behaviour model (primary, 3-class)}
\begin{center}
\begin{tabular}{lcccc}
\toprule
\textbf{Metric} & \textbf{Casual} & \textbf{At-Risk} & \textbf{Addicted} & \textbf{Overall} \\
\midrule
Precision & 0.943 & 0.916 & 0.913 & --- \\
Recall    & 0.938 & 0.900 & 0.948 & --- \\
F1        & 0.940 & 0.908 & 0.930 & --- \\
\midrule
Accuracy  & \multicolumn{4}{c}{\textbf{0.9243}} \\
Macro-F1  & \multicolumn{4}{c}{0.9261} \\
\bottomrule
\end{tabular}
\end{center}
Errors only ever fall into an \emph{adjacent} band (never Casual$\leftrightarrow$
Addicted). Probability calibration improves the multiclass Brier score from
$0.1254$ (uncalibrated) to $0.1213$ (isotonic), so reported confidences are
trustworthy. The accuracy is on a synthetic held-out split; the value of the result
is the calibrated, train/serve-aligned pipeline, not the absolute number.

\subsection{Chat toxicity model (honest, two views)}
\begin{center}
\begin{tabular}{lccc}
\toprule
\textbf{Evaluation set} & \textbf{Precision (toxic)} & \textbf{Recall (toxic)} & \textbf{Note} \\
\midrule
Balanced held-out & 0.887 & 0.931 & optimistic view \\
Realistic stream @0.50 & 0.048 & 0.931 & 0.7\% toxic base rate \\
Realistic stream @0.75 (alert) & 0.074 & 0.663 & operating threshold \\
\bottomrule
\end{tabular}
\end{center}
This is reported deliberately: on a realistic, rare-positive stream the model is
high-recall but low-precision (it over-flags), which is why the parent-facing alert
threshold is held high (0.75) and the UI labels borderline messages as ``borderline''
rather than ``toxic''. The genuine fix is real gaming-chat data, not a different
classifier --- a point we make explicit rather than hide behind the balanced number.

\subsection{Voice emotion model (synthetic)}
On a held-out split of synthetic acoustic features the model reaches $0.9925$
accuracy and $0.9925$ macro-F1 across angry / excited / frustrated / neutral. We
state plainly that this is \emph{internal validity on synthetic features}; real
microphone accuracy is unproven, which is why voice carries a corroborating weight
and is the first candidate for real-audio retraining.

\subsection{System and engineering results}
\begin{itemize}[leftmargin=1.4em]
  \item Production-deployed cloud backend (HTTPS, token auth, consent, durable
        managed Postgres); both apps build as signed release APKs.
  \item A layered test program, all passing: a 35-test API suite (isolated database),
        a \textbf{56-check functional sweep run in production auth mode} with real
        parent/child tokens (registration and family joins, role guards, the full
        nudge lifecycle, a real WAV upload through the silence floor and late
        re-score, the watchdog's quiet hours, feedback re-rating, parent-controlled
        deletion), and a \textbf{25-check end-to-end probe of the live cloud
        deployment}. Both apps pass Android Lint with zero errors.
  \item Robustness under load on free-tier hosting: pooled, probe-validated connections
        to the serverless Postgres (replacing per-request TLS handshakes; connections
        killed by scale-to-zero are detected and replaced transparently), per-client
        rate limiting behind the reverse proxy (previously all users shared one
        per-proxy quota --- the cause of earlier failures under load), memoised SHAP
        explanations and a throttled anomaly scan on the 30-second polling path, a
        16\,MB request-body cap, and threaded single-worker serving sized to the
        512\,MB instance. Verified with a concurrent smoke test (288 mixed requests
        across 24 threads, zero errors).
  \item Graceful degradation when a modality is absent, a self-healing stale-session
        guard, and an on-device offline retry queue for chat captured without
        connectivity.
\end{itemize}

\subsection{Real-device validation (and what it caught)}
Beyond the automated layers, the complete system was exercised on a physical Android~16
handset against the live cloud deployment: registration $\rightarrow$ consent
$\rightarrow$ the guided permission chain $\rightarrow$ \emph{automatic} session
detection on a real game (Roblox, no manual start) $\rightarrow$ live microphone
capture with on-device Vosk transcription (voice events and speech-to-text lines
arriving server-side and feeding the fusion) $\rightarrow$ automatic session end with a
scored prediction $\rightarrow$ the parent-side flows. This pass surfaced three
production-class defects that no amount of local or API-level testing had caught ---
each was fixed and is now regression-guarded:
\begin{enumerate}[leftmargin=1.4em]
  \item \textbf{Launch crash on modern Android.} Reading the enabled-keyboards system
        setting throws a \texttt{SecurityException} for apps targeting SDK~$>$~33, so
        the permission-chain check crashed the Home screen on Android~12+ devices; the
        check now uses the supported \texttt{InputMethodManager} API, and the
        remaining system-settings reads are wrapped to degrade to a re-prompt rather
        than a crash.
  \item \textbf{Out-of-memory under live voice load.} Real gameplay streams a voice
        segment every $\sim$10 seconds; several analysed concurrently pushed the
        512\,MB instance over its limit and triggered a platform restart. Fixed by the
        audio-analysis semaphore, burst-collapsing the per-segment re-score, more
        frequent worker recycling, and single-threaded BLAS (\S3.2).
  \item \textbf{A notification feedback loop.} The Parent dashboard (which shows the
        \emph{day roll-up} band) and the alert poller (which compares the
        \emph{latest-session} band) both wrote the same ``last seen risk'' state; near
        a band boundary the two disagreed, so every poll saw a ``change'' and re-fired
        the risk-change notification indefinitely. The poller is now the single
        writer, with a per-level re-notify cooldown absorbing genuine boundary
        flapping.
\end{enumerate}
We report these candidly: they are precisely the class of defect that only
real-device, real-network operation reveals, and finding and fixing them --- rather
than not looking --- is part of the production-readiness this paper claims.

\subsection{Summary of features delivered}
\begin{center}
\begin{tabularx}{\linewidth}{lX}
\toprule
\textbf{Area} & \textbf{Delivered} \\
\midrule
Detection & Any-game detection (UsageStats + category + parent overrides), session lifecycle with grace/ancillary handling \\
Capture & Custom-keyboard (IME) + Accessibility typed-chat (canvas games included), on-device Vosk STT, voice features with silence floor, real behavioural metrics, offline chat retry queue \\
ML & Calibrated RF behaviour + SHAP, LogReg/TF-IDF chat, GBM voice, availability-weighted fusion, valence--arousal voice/text fusion, genre weighting, late-voice re-score \\
Credibility & Probability calibration, anomaly detection (date-keyed), observation mode, honest model card, child-local timezone for hour-of-day signals \\
Parent value & Per-day weighted risk + ``why'' (SHAP), live ``playing now'' + monitoring-health strip, 14-day trend, recently/most-played games, real-time alerts with friendly ages + feedback agreement rate, weekly report (true weekly numbers) + PDF, limits, send-nudge-to-child, switch child without re-login \\
Child wellbeing & Mira check-in (typing indicator, double-send guard), gamified daily check-in + streak, reflections, breathing, break/limit nudges, today-vs-goal progress \\
Reliability/anti-tamper & Heartbeat watchdog (uninstall/kill/offline alert, child-local quiet hours), optional Device Admin (uninstall block + instant disable alert), battery-optimisation exemption, double-start-safe services, parent-PIN lock on logout/settings (server-verified, constant-time), parent-role-only PIN change + data deletion, logout \emph{and} login alerts with honest monitoring status, self-restart + boot resume, multi-guardian FCM push (lazy-init) \\
Privacy/Security & Token auth + role checks, peppered PIN hashes, HTTPS-only, encrypted token store, consent (versioned), raw-audio deletion, retention purge, 16\,MB body cap, family code always recoverable in PIN-gated Settings \\
Engineering & Dual-dialect DB with pooled validated Postgres connections, per-client rate limiting, hot-path caching/throttling + audio-concurrency cap, signed APKs, layered tests (35 API + 56 production-mode + 25 live-cloud) + concurrency smoke, real-device validated, lint-clean, deployed + durable DB \\
\bottomrule
\end{tabularx}
\end{center}

% =====================================================================
\section{Conclusion}
% =====================================================================
We built and deployed a multimodal gaming-addiction \emph{screening} platform that
moves beyond raw screen-time to measured behaviour, chat toxicity and voice emotion,
fused into a calibrated, explainable risk band and surfaced to parents as awareness
rather than punishment. The engineering is real: native Android apps, a secured cloud
backend, a durable database, signed releases, tamper-resistant background monitoring with
real-time multi-guardian alerts, a layered passing test program (API suite,
production-mode functional sweep, live-cloud probe), and end-to-end validation on real
hardware against the live deployment. Our central
methodological commitments --- train/serve feature alignment, probability
calibration, SHAP explanations, availability-weighted fusion, and an honest account
of synthetic-data limitations --- make the system's outputs trustworthy and
defensible. Most importantly, the parent-feedback loop turns the platform from a
frozen, synthetically-trained model into one that can learn from real verdicts,
which is the credible path from a working prototype to a validated product.

% =====================================================================
\section{Future Work}
% =====================================================================
We group future enhancements by how feasible they are for the project team to build,
distinguishing near-term engineering work from directions that need external data,
partners, or hardware.

\textbf{Tier 1 --- Immediate, high-value extension (engineering only).}
The one near-term enhancement genuinely worth building with the existing codebase and free
tools --- one that adds visible value and can be implemented directly --- is:
\begin{itemize}[leftmargin=1.4em]
  \item \textbf{Real-audio voice model + VAD.} Retrain the voice model on free, real
        labelled speech (RAVDESS~\cite{ravdess2018}) and add full voice-activity detection
        (\texttt{webrtcvad}). The served pipeline already applies an RMS silence floor so
        silent segments are not classified as emotion; real VAD additionally rejects
        \emph{game audio and background speech}, which an energy floor cannot. Voice
        is the weakest modality and the \emph{only} one trained on synthetic data, so this
        is the single change that most strengthens the system's credibility; the dataset
        and tooling are freely available and the retraining pipeline already exists.
\end{itemize}

\textbf{Tier 2 --- Data-dependent improvements.}
The code for these is tractable, but their value depends on accumulating real usage
data the prototype does not yet have:
\begin{itemize}[leftmargin=1.4em]
  \item \textbf{Feedback-driven retraining.} The feedback loop already records parent
        ``accurate / false-alarm'' labels; a retrain script can ingest them to tune the
        alert threshold and chat boundary, then learn ensemble weights. The pipeline is
        buildable now, but meaningful accuracy gains need a real pilot at scale --- this
        is the single change that converts synthetic accuracy into validated performance.
  \item \textbf{Per-user personalisation.} Flag deviations from a child's personal
        baseline rather than only global thresholds (a professional player's ``normal''
        differs from a student's). Straightforward to code; needs weeks of real
        per-child history to calibrate.
  \item \textbf{Cross-modal interaction features.} Add features such as ``angry voice
        while late-night'' to a meta-model for sharper escalation detection. Easy to
        engineer; needs real labels to validate they help.
  \item \textbf{Multilingual / code-mixed chat.} Real gaming chat among the target users
        is heavily code-mixed (e.g.\ Hinglish and romanised regional text), which the
        current English TF-IDF model misses (cf.\ the multilingual sentiment work in our
        literature survey). Swapping in a pretrained multilingual model is feasible; the
        blocker is labelled code-mixed gaming-chat data.
\end{itemize}

\textbf{Tier 3 --- Longer-term research directions.}
These require partners, hardware, or scale beyond a capstone window and are described as
direction rather than as work we can execute alone:
\begin{itemize}[leftmargin=1.4em]
  \item \textbf{Expert / clinical validation.} Validate the risk bands against an
        established instrument (e.g.\ an Internet Gaming Disorder scale) with a
        counsellor or psychologist and real subjects --- the honest path from a screening
        signal to a clinically-meaningful one. This is a supervised research study (with
        ethics approval), not a coding task.
  \item \textbf{Sleep / wearable corroboration.} Corroborate late-night play with sleep
        data from a wearable to strengthen the sleep-disruption signal. Needs hardware
        and real users; the app already logs screen-off / device-idle patterns as a
        lightweight proxy in the meantime.
\end{itemize}

We deliberately \emph{do not} pursue on-device TFLite conversion, transformer/CNN
swaps, or federated learning at this stage: on synthetic data they add complexity and
deployment risk without proven benefit, and the honest, calibrated classical ensemble
is both defensible and shippable today.

\vspace{1.5em}
\hrule
\vspace{0.5em}
{\small \textit{All figures and metrics in this paper reflect the system as
implemented and deployed; model metrics are the current values recorded in the
project's model metadata. The system is a screening and awareness aid, not a
clinical diagnostic instrument.}}

% =====================================================================
\begin{thebibliography}{99}
\small
\bibitem{huang2024} Y. Huang, R. Wu, Y. Huang, Y. Xiang, and W. Zhou,
``Investigating mechanisms of Internet Gaming Disorder and developing intelligent
monitoring models using artificial intelligence technologies,'' \emph{BMC Public
Health}, 2024.
\bibitem{jiang2024} Z. Jiang, ``A comparative analysis of contact and non-contact
approaches using machine learning for gaming disorder detection,'' in \emph{Proc.\ Int.\
Conf.\ Modern Logistics and Supply Chain Management (MLSCM)}, 2024.
\bibitem{ergin2025} D. A. Ergin and C. A. Essau, ``Family factors and Internet Gaming
Disorder among adolescents: A systematic review,'' \emph{Int.\ J.\ Developmental
Science}, 2025.
\bibitem{kaya2023} M. Kaya, A. Akg\"ul, and \"O. Kaya, ``Online gaming addiction and
basic psychological needs: A study among adolescents,'' \emph{Int.\ J.\ Mental Health and
Addiction}, 2023.
\bibitem{strojny2024} P. Strojny, K. Kapela, N. Lipp, and S. Sikstr\"om, ``Use of four
open-ended text responses to help identify people at risk of gaming disorder,'' \emph{JMIR
Serious Games}, vol.\ 12, e56663, 2024.
\bibitem{kina2024} E. Kina, U. G\"uven, and B. Kara, ``Multilingual sentiment analysis
for mobile gaming: A comparative study of machine learning and hybrid deep learning
approaches,'' \emph{SETSCI Conference Proceedings}, 2024.
\bibitem{tyagi2024} S. Tyagi, A. S. Ansari, and A. Raina, ``Semantic speech analysis
using machine learning and multimodal data processing,'' \emph{Multimedia Tools and
Applications}, 2024.
\bibitem{schuller2013} B. Schuller \emph{et al.}, ``Paralinguistics in speech and
language --- State-of-the-art and the challenge,'' \emph{Computer Speech \& Language},
2013.
\bibitem{chakraborty2025} S. Chakraborty, ``Integrating machine learning on mobile
applications using local data to personalize the experience for the customer,''
ResearchGate Preprint, 2025.
\bibitem{iftikhar2021} A. Iftikhar, M. Mobin, and M. Sarwar, ``Designing parental
monitoring and control technology,'' in \emph{Proc.\ INTERACT}, 2021.
\bibitem{bark2025} Bark Technologies, Inc., ``Bark parental control app --- AI-powered
monitoring for texts, apps, and online activity,'' SafeWise Review, 2025.
\bibitem{dsm5} American Psychiatric Association, \emph{Diagnostic and Statistical Manual
of Mental Disorders (DSM-5)}, 5th ed., Internet Gaming Disorder (Section III), 2013.
\bibitem{icd11} World Health Organization, \emph{International Classification of Diseases
11th Revision (ICD-11)} --- Gaming disorder (6C51), 2019.
\bibitem{shap2017} S. M. Lundberg and S.-I. Lee, ``A unified approach to interpreting
model predictions,'' in \emph{Advances in Neural Information Processing Systems
(NeurIPS)}, 2017.
\bibitem{ravdess2018} S. R. Livingstone and F. A. Russo, ``The Ryerson Audio-Visual
Database of Emotional Speech and Song (RAVDESS),'' \emph{PLoS ONE}, vol.\ 13, no.\ 5,
2018.
\end{thebibliography}

\end{document}
